---
title: "BC big game harvesting"
output: html_notebook
---

<!-- 
Execute a chunk: click the *Run* button within the chunk or place your cursor inside it and press *Ctrl+Shift+Enter*.

Add a new chunk:  click the *Insert Chunk* button on the toolbar or press *Ctrl+Alt+I*. 

https://stackoverflow.com/a/4714080
- @ accesses a slot in an S4 object like an object in spatialPolygons class.  Find slot names using slotNames(spPolygonObjectName).  Find contents of a slot using slot(spPolygonObjectName, "slotName").
- $ accesses a named variable in an S4 object or a dataframe or a list
-->

This notebook analyzes data on big game harvesting in British Columbia, Canada.  It uses the following publicly-available data from the Government of BC:

* Hunting kills dataset covering 1976--2018 from: https://catalogue.data.gov.bc.ca/dataset/big-game-harvest-statistics-1976-2018
  - I used the .xlsx version of the dataset rather than the .csv version, because the csv had half the header text cut off and does not have descriptions of field codes
  - The BC government has a Shiny app for this data at:  https://kootenaywildlife.shinyapps.io/BCHarvestData/
* Annual population of BC from:  https://www2.gov.bc.ca/gov/content/data/statistics/people-population-community/population/population-estimates
* Hunting licence sales from:
https://catalogue.data.gov.bc.ca/dataset/hunting-sales-statistics-2005-2019
* BC Wildlife management units from:
https://catalogue.data.gov.bc.ca/dataset/wildlife-management-units

Some possibilities for data on species populations (incomplete):

* GRIZZLY BEAR:  bc_grizzly_population_estimates_2015_and_2018_by_gbpu_population_sub_units from https://catalogue.data.gov.bc.ca/dataset/grizzly-bear-population-estimates
* MOOSE:  Kuzyk -- 2016 -- Provincial population and harvest estimates of moose in British Columbia
* ELK:  here, but no public access:  https://catalogue.data.gov.bc.ca/dataset/elk-cervus-elaphus-populations-scale-1-250-000
* MULE DEER:  maybe in the aerial survey of large mammals, but access restricted:  https://catalogue.data.gov.bc.ca/dataset/aerial-survey-large-mammals-1994-1997
* CARIBOU:  herd locations:  https://catalogue.data.gov.bc.ca/dataset/caribou-herd-locations-for-bc
* WILDLIFE SPECIES INVENTORY:  https://catalogue.data.gov.bc.ca/dataset/wildlife-species-inventory-results-by-area-all
* Francis Iredale (via Jim Gnapp) suggests for species population data:
  - http://a100.gov.bc.ca/pub/siwe/search_reset.do
  - https://a100.gov.bc.ca/pub/clir/enterCLIR.do?redirect=Y
  
Other possible resources:

* http://www.env.gov.bc.ca/fw/wildlife/management-issues/docs/kootenay_mule_deer_faq.pdf
* try getting population data from here:  http://a100.gov.bc.ca/pub/siwe/search_reset.do
* BC hunter expenditure survey:  http://www.env.gov.bc.ca/fw/wildlife/bc_hunt_value.html
* BC harvest allocation policies:  http://www.env.gov.bc.ca/fw/wildlife/harvest_alloc/
* region and wmu info:  https://www2.gov.bc.ca/gov/content/sports-culture/recreation/fishing-hunting/hunting/regulations-synopsis
* BC regions map:  https://governmentofbc.maps.arcgis.com/home/webmap/viewer.html
* BC data sets:  catalogue.data.gov.bc.ca/dataset
* BC game management zones:  https://catalogue.data.gov.bc.ca/dataset/game-management-zones

Author:  Heather More
Started:  2020-11-24 

```{r}
# INPUTS
# Which species to plot
speciesToShow <- c("DEMU", "BEAB")#"ELK", "DEMU", "DEWT", "MOOS", "BEAB", "WOLF", "CARI", "GOAT", "SHEE", "BEAG", "COUG")
# Which WMU to plot -- as number, without dash, single digit as 0X
wmuToShow <- c(332)
# Which hunting region to plot -- as number, or as string if 7A or 7B
huntingRegionToShow <- c("7A")
# Choose to filter by province totals, hunting regions, or wmus
# Options:  "province", "huntingRegion", "wmu"
filterBy <- "huntingRegion"
```

```{r}
# PARAMETERS

# Plot options
map_strokeWeight <- 2
map_strokeOpacity <- 1
map_fillOpacity <- 0.4
map_hoverColour <- "red"
map_hoverStrokeWeight <- 2
map_staticLabelColour <- "grey"
```

```{r}
# LOAD PACKAGES
library(readxl) # read Excel spreadsheets
library(tidyverse) # data manipulation
library(janitor) # data cleaning
library(leaflet)  # mapping
library(geojsonio)  # read geojson files
library(rmapshaper)  # simplify areas on maps
library(rgeos)  # functions for handling geographic regions
library(geosphere)  # functions for handling sperical regions (i.e. lat/long coordinages)
library(RColorBrewer)  # colour schemes
library(htmltools)  # HTML formatting for map labels

# CUSTOM FUNCTIONS
# Convert names of colours to hex
source("GetColourHex.R")
# Custom functions for this code
source("CustomFunctions.R")

# INPUT FILES
# Directory containing data
data_dir = "D:/Documents/Code/BC big game harvesting/data"
# Historical hunting kills
data_huntingKills <- list(
  "path" = data_dir,
  "fileName" = "big-game-harvest-statistics-1976-2018.xlsx",
  "sheetName" = "BGHS",
  "columnTypes" = c("numeric",  # hunt year
                     "text",  # species
                     "text",  # CI
                     "numeric",  # prov flag
                     "text",  # WMU
                     "text",  # region
                     rep("numeric", 14)  # remaining columns
                    )
)
# Historical population of BC
data_bcPopulation <- list(
  "path" = data_dir,
  "fileName" = "pop_bc_annual_estimates_clean.xlsx",
  "sheetName" = "data"
)
# Lists of WMUs and their corresponding hunting regions
data_wmuHuntingRegions <- list(
  "path" = data_dir,
  "fileName" = "conversions_hunting-region-wmus.xlsx",
  "sheetName" = "wmu_data",
  "columnTypes" = c("text",  # WMU id as string
                    "numeric",  # WMU id as int
                    "text",  # hunting region
                    "text"  # notes
                    )
)
# Lists of hunting region names
data_regionNames <- list(
  "path" = data_dir,
  "fileName" = "conversions_hunting-region-wmus.xlsx",
  "sheetName" = "region_names",
  "columnTypes" = c("text",  # region id as string
                    "text"  # region name as string
                    )
)
# List of species included in the hunting dataset and their attributes
data_species <- list(
  "path" = data_dir,
  "fileName" = "conversions_species.xlsx",
  "sheetName" = "data"
)
# Map of WMUs
map_wmu <- list(
  "path" = data_dir,
  "fileName" = "wmu-map/WAA_WILDLIFE_MGMT_UNITS_SVW.geojson"
)

# LINKS
# Site to preview different basemaps:  http://leaflet-extras.github.io/leaflet-providers/preview/index.html
#tilesURL <- "http://server.arcgisonline.com/ArcGIS/rest/services/Canvas/World_Light_Gray_Base/MapServer/tile/{z}/{y}/{x}"
tilesURL <- "https://server.arcgisonline.com/ArcGIS/rest/services/World_Topo_Map/MapServer/tile/{z}/{y}/{x}"
```

```{r}
# LOAD DATA

# Hunting kills
df_huntingKills <- ReadExcel(data_huntingKills)

# BC population
df_pop <- ReadExcel(data_bcPopulation)

# Conversions between WMU numbers and hunting regions
df_wmuRegions <- ReadExcel(data_wmuHuntingRegions)

# Hunting region names
df_regionNames <- ReadExcel(data_regionNames)

# Species names and characteristics
df_species <- ReadExcel(data_species)

# Map of wildlife management units.
# https://rstudio.github.io/leaflet/json.html
# Simplify boundaries so they render faster -- size goes from 4,535,784 to 963,312 bytes
sp_wmus <- geojsonio::geojson_read(file.path(map_wmu[["path"]], map_wmu[["fileName"]], 
                           fsep = .Platform$file.sep), what = "sp") %>% 
  rmapshaper::ms_simplify()
# Try saving reduced-size map to lower RShiny load time
# https://towardsdatascience.com/interactive-covid19-report-with-rmarkdown-plotly-leaflet-and-shiny-c6a716af7d9b
# saveRDS(world_geojson, "countries_map.rds")

```
```{r}
# FUNCTIONS

# Look up full name of species based on code
# Input is species code as string, e.g. "BEAB"
# Output is species name as string, e.g. "Black bear"
GetSpeciesName <- function(speciesCode) {
  species_name <- df_species %>% 
    filter(code==speciesCode) %>% 
    pull(name)
  return(species_name)
}

# Find colour corresponding to species
# Input is species code as string, e.g. "BEAB"
# Output is colour as hex string, e.g. "#000000"
GetSpeciesColour <- function(speciesCode) {
  species_colour <- df_species %>% 
    filter(code==speciesCode) %>% 
    pull(colour) %>% 
    GetColourHex()
  return(species_colour)
}

# Rename entries in residency columns to "resident" or "non-resident" when they contain other information
# Input:  x is a column from a dataframe, as a list; get it by df$x
# Output:  renamed_x is a list that can be put back into a dataframe
RenameResidentEntries <- function(x) {
  x[str_detect(x, "non")] <- "non-resident"
  x[str_detect(x, "non", negate = TRUE)] <- "resident"
  return(x)
}

# Find hunting region of a WMU
# Input:  WMU number as string
# Output:  Hunting region number as string
GetRegionCode <- function(wmuID) {
  regionString <- df_wmuRegions %>% 
    filter(wmu_string==wmuID) %>% 
    pull(region)  # return column as a vector
  return(regionString)
}
# Get name of hunting region from region number
# Input:  Hunting region number as string
# Output:  Hunting region name as string
GetRegionName <- function(regionID) {
  regionName <- df_regionNames %>% 
    filter(region==regionID) %>% 
    pull(name)  # return column as a vector
  return(regionName)
}

# Filter hunting data based on criteria.
# Using a function allows all the manipulation to happen in one go, so that in Shiny we will only have to define the reactive dataframe once.  There must be other ways to do this but I am not sure what they are.  It's also (maybe?) nicer than having a giant if elseif statement in the middle of everything.
# df_huntingKills$prov_flag
#     9 for whole-province data
#     0 otherwise
# df_huntingKills$wmu
#     999 for whole-province data
#     X00 for all of region X
#     7A or 7B for all of region 7A or 7B
# Inputs:
# - input dataframe
# - a single string denoting how to filter the input dataframe -- currently "province", "huntingRegion", "wmu"
# Output:
# - filtered dataframe
FilterByRegion <- function(df_in, filterCriteria) {
  if (filterCriteria == "province") {
    df_out <- df_in %>%
      filter(prov_flag == 9)
  } else if (filterCriteria == "huntingRegion") {
    if (is.character(huntingRegionToShow)) {
      df_out <- df_in %>%
        filter(wmu == huntingRegionToShow)
    } else if (is.numeric(huntingRegionToShow)) {
      df_out <- df_in %>%
        filter(wmu == (100 * huntingRegionToShow + 99))
    }
  } else if (filterCriteria == "wmu") {
    df_out <- df_in %>%
      filter(wmu == wmuToShow)
  }
}

```

```{r}
# INVESTIGATE DATA
print(paste("Hunt years:  ", 
            paste(min(df_huntingKills$hunt_year)), 
            "--", 
            paste(max(df_huntingKills$hunt_year)), 
            collapse = ''
            )
      )
print(paste('Species:  ', 
            paste(lapply(unique(df_huntingKills$species), GetSpeciesName), 
                  collapse = ', ')
            )
      )
print(paste('WMUs:  ', 
            paste(length(unique(df_huntingKills$wmu)))
            )
      )
print(paste('Regions:  ', 
            paste(length(unique(df_huntingKills$region)))
            )
      )
```

```{r}
# FILTER AND ADD OTHER VARIABLES
# Do this in one go for each dataframe, so that in Shiny it's easier to make the dataframes reactive

# Filter population data so years match hunting data
df_pop_years <- df_pop %>% 
  filter(year >= min(df_huntingKills$hunt_year) 
         & year <= max(df_huntingKills$hunt_year)) %>% 
  rename(population_human = population)

# Filter hunting data
df_filtered <- df_huntingKills %>% 
  # Show selected species
  filter(species %in% speciesToShow) %>% 
  # Show selected region
  FilterByRegion(., filterBy) %>% 
  # Add population data
  left_join(df_pop_years, by = c("hunt_year" = "year")) %>% 
  # Compute new metrics
  rowwise() %>% # sum over each row:  https://stackoverflow.com/a/33806717
  mutate(
    total_kills = sum(resident_kills, non_resident_kills, na.rm=TRUE),
    total_hunters = sum(resident_hunters, non_resident_hunters, na.rm=TRUE),
    total_days = sum(resident_days, non_resident_days, na.rm=TRUE)
    )

# Compute totals for residents and non-residents over all species for each year
df_year_totals <- df_filtered %>% 
  # Compute totals per year
  group_by(hunt_year) %>% 
  summarise(
    across(resident_hunters, sum, na.rm=TRUE),
    across(non_resident_hunters, sum, na.rm=TRUE),
    across(resident_kills, sum, na.rm=TRUE),
    across(non_resident_kills, sum, na.rm=TRUE)
  ) %>% 
  # Add population of province in each year
  left_join(df_pop_years, by = c("hunt_year" = "year")) %>% 
  # Pivot # hunters
  pivot_longer(cols = contains("hunters"),
               names_to = "hunter_residency", 
               values_to = "n_hunters"
  ) %>% 
  # Pivot # kills
  pivot_longer(cols = contains("kills"),
               names_to = "kill_residency", 
               values_to = "n_kills"
  )  %>% 
  # Rename entries in resident columns to "resident" and "non-resident"
  mutate(hunter_residency = RenameResidentEntries(hunter_residency),
         kill_residency = RenameResidentEntries(kill_residency)
  ) %>% 
  # Remove rows where hunter residency and kill residency are different -- these are an artifact of doing two pivots
  filter(hunter_residency == kill_residency) %>% 
  # Remove redundant residency column and rename remaining column
  mutate(kill_residency = NULL) %>% 
  rename(residency = hunter_residency)
```
```{r}
# ADD VARIABLES TO MAP DATA AND GENERATE MAP OF HUNTING REGIONS
#
# WMU codes in hunting data are e.g. 101, 115
#   - codes ending in 00 denote values from that hunting region but unknown WMU
#   - codes ending in 99 denote totals for that hunting region
#   - code 999 is province-wide total
# WMU codes in geojson are e.g. 1-1, 1-15

# Check total areas add to area of BC (944,735 km^2)
# Note that the geojson areas include a bunch of water so we expect a higher estimate
#testarea <- sum(wmus$FEATURE_AREA_SQM/(1000^2))

# ADD DATA TO WMU MAP

# Add fields to separate region 7 into 7A and 7B (region code and region name)
sp_wmus@data$HUNTING_REGION_ID <- unlist(lapply(sp_wmus$WILDLIFE_MGMT_UNIT_ID, 
                                          GetRegionCode))
sp_wmus@data$HUNTING_REGION_NAME <- unlist(lapply(sp_wmus$HUNTING_REGION_ID, 
                                          GetRegionName))

# GENERATE HUNTING REGION MAP

# Group WMU boundaries into hunting regions
# https://gis.stackexchange.com/a/225731
sp_huntingRegions <- gUnaryUnion(sp_wmus, id = sp_wmus@data$HUNTING_REGION_ID)

# ADD DATA TO HUNTING REGION MAP

# Add dataframe of hunting region IDs to SpatialPolygons object; this turns it into a SpatialPolygonsDataFrame and allows us to add data for each region
# - ID of each polygon is hunting region.
# - dataframe column names will be converted to SpatialPolygons tags
# - dataframe row names must be the same as polygon IDs
# Resources:
# - https://willchernoff.com/2012/10/08/add-attribute-data-to-object-of-class-spatialpolygonsdataframe-in-r/
# - to get polygon IDs:  https://r.789695.n4.nabble.com/How-extract-the-names-of-ID-in-SpatialPolygons-object-td790614.html
hunting_regions_data <- tibble(
  sapply(slot(sp_huntingRegions, "polygons"), function(x) slot(x, "ID"))
  )
colnames(hunting_regions_data) <- c("HUNTING_REGION_ID")
rownames(hunting_regions_data) <- hunting_regions_data$HUNTING_REGION_ID
# Combind SpatialPolygons object and dataframe to generate SpatialPolygonsDataFrame
sp_huntingRegions <- SpatialPolygonsDataFrame(sp_huntingRegions, 
                                            hunting_regions_data
                                            )

# NAME of each hunting region
sp_huntingRegions@data$HUNTING_REGION_NAME <- unlist(lapply(sp_huntingRegions$HUNTING_REGION_ID, 
                                          GetRegionName))
# AREA of each hunting region (km^2)
# Note that "area" in a polygon class is NOT the projected area, but rather a planar area measurement used to ensure smaller polygons are plotted after larger ones:  https://www.rdocumentation.org/packages/sp/versions/1.4-4/topics/Polygons-class
# Instead, from the rgeos package use gArea() (for projected coordinates) or from the geosphere package use areaPolygon() (for lat-long coordinates) (check the +proj= part of the SpatialPolygon file to find out which you are using):  https://stackoverflow.com/a/8708828
sp_huntingRegions@data$HUNTING_REGION_AREA <- areaPolygon(sp_huntingRegions)/(1000^2)

```

```{r}
# PLOTTING FUNCTIONS

# Theme
UseTheme <- function() {
    theme_light()
}

# Line and point styles
FormatLines <- function() {
  geom_line()
}
FormatPoints <- function() {
  geom_point(size=2, alpha=0.7)
}

# Use species-specific colours in line plot
UseSpeciesColours <- function() {
     scale_colour_manual(
       name = "Species",  # legend title
       values = paste(lapply(speciesToShow, GetSpeciesColour)),  # colour hex values
       breaks = speciesToShow,  # species codes corresponding to each hex value
       labels = paste(lapply(speciesToShow, GetSpeciesName))  # species names for legend
     )
}

# Text identifying year range
YearRangeAsText <- function() {
   paste("from", 
         paste(min(df_filtered$hunt_year)), 
         "to", 
         paste(max(df_filtered$hunt_year))
   )
}

# Generate colour palette for maps
# Show possible colour schemes
#display.brewer.all()
mapColours <- colorFactor(
  palette = brewer.pal(length(unique(sp_wmus$HUNTING_REGION_ID)), "Set3"),
  levels = unique(sp_wmus$HUNTING_REGION_ID)
  )

# Add polygons with consistent style
# Inputs:
# - Leaflet map (just use .)
# - SpatialPolygonDataFrame with polygons and data field containing colour variable
# - name of variable to colour polygons by; must be in data field of SpatialPolygonDataFrame
# - text for labels on hover; can include HTML
# Output:  plots polygons with popup labels on the leaflet map
PlotPolygons <- function(leafletObject, inputData, colourVariableName, labelText) {
  leafletObject %>% 
    addPolygons(data = inputData,
                stroke = TRUE,
                weight = map_strokeWeight,
                opacity = map_strokeOpacity,
                fillOpacity = map_fillOpacity,
                color = ~mapColours(inputData[[colourVariableName]]),
                # Text that pops up on hover:  https://stackoverflow.com/a/43155126
                label = lapply(labelText, htmltools::HTML),
                labelOptions = labelOptions(direction = "top"),
                # Polygon appearance on hover
                highlight = highlightOptions(color = map_hoverColour,
                                             fillColor = map_hoverColour,
                                             weight = map_hoverStrokeWeight,
                                             bringToFront = TRUE
                )
  )
}
# Add static text to label polygons
# Inputs:
# - Leaflet map (just use .)
# - Dataframe containing columns:
#   - "long" = longitude coordinate of each label
#   - "lat" = latitude coordinate of each label
#   - "labelVariableName" = text to show for each label; can include HTML
# Output:  none, just applies labels
AddStaticLabels <- function(leafletObject, labelData, labelVariableName) {
  leafletObject %>% 
    addLabelOnlyMarkers(data = labelData,
                      lng = ~long, 
                      lat = ~lat, 
                      label = labelData[[labelVariableName]],
                      labelOptions = labelOptions(
                        noHide = TRUE, 
                        direction = 'auto', 
                        textOnly = TRUE,
                        style = list(
                          "color" = map_staticLabelColour
                          )
                        )
                      )
}

```

```{r}
# Total kills
plot_prov_species <- ggplot(data = df_filtered) +
  aes(x = hunt_year, y = total_kills, color = species) +
  FormatLines() +
  FormatPoints() +
  UseTheme() +
  UseSpeciesColours() +
  labs(x = paste("Year"),
       y = paste("Total # kills"),
       title = paste("Total kills", YearRangeAsText())
  )
print(plot_prov_species)
```


```{r}
# Total hunters
plot_prov_hunters <- ggplot(data = df_filtered) +
  aes(x = hunt_year, y = total_hunters, color = species) +
  FormatLines() +
  FormatPoints() +
  UseTheme() +
  UseSpeciesColours() +
  labs(x = paste("Year"),
       y = paste("Total # hunters"),
       title = paste("Total hunters", YearRangeAsText())
  )
print(plot_prov_hunters)
```


```{r}
# Avg kills per hunter
plot_prov_killsperhunter <- ggplot(data = df_filtered) +
  aes(x = hunt_year, y = total_kills/total_hunters, color = species) +
  FormatLines() +
  FormatPoints() +
  UseTheme() +
  UseSpeciesColours() +
  labs(x = paste("Year"),
       y = paste("Avg kills per hunter"),
       title = paste("Average kills per hunter", YearRangeAsText())
  )
print(plot_prov_killsperhunter)
```
```{r}
# Total hunting days
plot_prov_totaldays <- ggplot(data = df_filtered) +
  aes(x = hunt_year, y = total_days, color = species) +
  FormatLines() +
  FormatPoints() +
  UseTheme() +
  UseSpeciesColours() +
  labs(x = paste("Year"),
       y = paste("Total # hunting days"),
       title = paste("Total # hunting days", YearRangeAsText())
  )
print(plot_prov_totaldays)
```

```{r}
# Avg hunting days per hunter
plot_prov_daysperhunter <- ggplot(data = df_filtered) +
  aes(x = hunt_year, y = total_days/total_hunters, color = species) +
  FormatLines() +
  FormatPoints() +
  UseTheme() +
  UseSpeciesColours() +
  labs(x = paste("Year"),
       y = paste("Avg hunt days per hunter"),
       title = paste("Average # hunting days per hunter", YearRangeAsText())
  )
print(plot_prov_daysperhunter)
```

```{r}
# Avg hunting days per kill
plot_prov_daysperhunter <- ggplot(data = df_filtered) +
  aes(x = hunt_year, y = total_days/total_kills, color = species) +
  FormatLines() +
  FormatPoints() +
  UseTheme() +
  UseSpeciesColours() +
  labs(x = paste("Year"),
       y = paste("Avg hunt days per kill"),
       title = paste("Average # hunting days per kill", YearRangeAsText())
  )
print(plot_prov_daysperhunter)
```



```{r}
# Resident hunters of each species
plot_prov_speciesres <- ggplot(data = df_filtered) +
  aes(x = hunt_year, y = resident_hunters, color = species) +
  FormatLines() +
  FormatPoints() +
  UseTheme() +
  UseSpeciesColours() +
  labs(x = paste("Year"),
       y = paste("# resident hunters"),
       title = paste("Resident hunters", YearRangeAsText())
  )
print(plot_prov_speciesres)
```

```{r}
# Non-resident hunters of each species
plot_prov_speciesnonres <- ggplot(data = df_filtered) +
  aes(x = hunt_year, y = non_resident_hunters, color = species) +
  FormatLines() +
  FormatPoints() +
  UseTheme() +
  UseSpeciesColours() +
  labs(x = paste("Year"),
       y = paste("# non-resident hunters"),
       title = paste("Non-resident hunters", YearRangeAsText())
  )
print(plot_prov_speciesnonres)
```

```{r}
# Resident hunters of each species, normalized to BC population
plot_prov_speciesresfrac <- ggplot(data = df_filtered) +
  aes(x = hunt_year, y = resident_hunters/population_human, color = species) +
  FormatLines() +
  FormatPoints() +
  UseTheme() +
  UseSpeciesColours() +
  labs(x = paste("Year"),
       y = paste("# resident hunters/total BC population"),
       title = paste("Resident hunters as fraction of BC population", YearRangeAsText())
  )
print(plot_prov_speciesresfrac)
```

```{r}
# Total resident and non-resident hunters
plot_prov_totalresnonres <- ggplot(data = df_year_totals) +
  aes(x = hunt_year, y = n_hunters, color = residency) +
  FormatLines() +
  FormatPoints() +
  UseTheme() +
  labs(x = paste("Year"),
       y = paste("# hunters"),
       title = paste("Total resident and non-resident hunters", YearRangeAsText())
  )
print(plot_prov_totalresnonres)
```



```{r}
plot_prov_totalresfrac <- ggplot(data = filter(df_year_totals, residency == "resident")) +
  aes(x = hunt_year, y = n_hunters/population_human) +
  FormatLines() +
  FormatPoints() +
  UseTheme() +
  labs(x = paste("Year"),
       y = paste("# hunters/BC population"),
       title = paste("Total resident hunters as a fraction of BC population", YearRangeAsText())
  )
print(plot_prov_totalresfrac)
```

```{r}
plot_bcpop <- ggplot(data = df_pop_years) +
  aes(x = year, y = population_human) +
  FormatLines() +
  FormatPoints() +
  UseTheme() +
  labs(x = paste("Year"),
       y = paste("BC population"),
       title = paste("BC population", YearRangeAsText())
  )
print(plot_bcpop)
```


```{r}
# Avg kills per hunter, separated by hunter residency
plot_prov_killsperhunter_resident <- ggplot(data = df_filtered) +
  aes(x = hunt_year, y = resident_kills/resident_hunters, color = species) +
  FormatLines() +
  FormatPoints() +
  UseTheme() +
  UseSpeciesColours() +
  labs(x = paste("Year"),
       y = paste("Avg kills per resident hunter"),
       title = paste("Average kills per resident hunter", YearRangeAsText())
  )
print(plot_prov_killsperhunter_resident)
```

```{r}
plot_prov_killsperhunter_nonresident <- ggplot(data = df_filtered) +
  aes(x = hunt_year, y = non_resident_kills/non_resident_hunters, color = species) +
  FormatLines() +
  FormatPoints() +
  UseTheme() +
  UseSpeciesColours() +
  labs(x = paste("Year"),
       y = paste("Avg kills per non-resident hunter"),
       title = paste("Average kills per non-resident hunter", YearRangeAsText())
  )
print(plot_prov_killsperhunter_nonresident)
```

```{r}
# Avg hunting days per kill, separated by hunter residency
plot_prov_daysperhunter_resident <- ggplot(data = df_filtered) +
  aes(x = hunt_year, y = resident_days/resident_kills, color = species) +
  FormatLines() +
  FormatPoints() +
  UseTheme() +
  UseSpeciesColours() +
  labs(x = paste("Year"),
       y = paste("Avg hunt days per kill"),
       title = paste("Average # hunting days per kill for resident hunters", YearRangeAsText())
  )
print(plot_prov_daysperhunter_resident)
```

```{r}
plot_prov_daysperhunter_nonresident <- ggplot(data = df_filtered) +
  aes(x = hunt_year, y = non_resident_days/non_resident_kills, color = species) +
  FormatLines() +
  FormatPoints() +
  UseTheme() +
  UseSpeciesColours() +
  labs(x = paste("Year"),
       y = paste("Avg hunt days per kill"),
       title = paste("Average # hunting days per kill for non-resident hunters", YearRangeAsText())
  )
print(plot_prov_daysperhunter_nonresident)
```

```{r}
# Total kills over all selected species for resident and non-resident hunters
plot_prov_totalkills <- ggplot(data = df_year_totals) +
  aes(x = hunt_year, y = n_kills, color = residency) +
  FormatLines() +
  FormatPoints() +
  UseTheme() +
  labs(x = paste("Year"),
       y = paste("Total kills"),
       title = paste("Total kills", YearRangeAsText())
  )
print(plot_prov_totalkills)
```

```{r}
# Map of WMUs

# PREPARE LABELS
# Labels to show on each WMU all the time
wmu_static_labels <- lapply(sp_wmus@polygons, function(x) {x@labpt}) %>% 
  unlist() %>% 
  matrix(., nrow = length(.), ncol = 2, byrow = TRUE) %>% 
  as_tibble() %>% 
  cbind(sp_wmus$WILDLIFE_MGMT_UNIT_ID)
colnames(wmu_static_labels) <- c("long", "lat", "WMU")
# Labels to show on each WMU when hovering
#https://stackoverflow.com/a/43155126
wmu_hover_labels <- lapply(seq(nrow(sp_wmus)),
                           function(x) {
                             paste(
                               paste(
                                 "<b>WMU # ", 
                                 sp_wmus$WILDLIFE_MGMT_UNIT_ID[x], 
                                 "</b>", 
                                 sep = ''
                                 ),
                               paste(
                                 format(
                                   round(as.numeric(sp_wmus$FEATURE_AREA_SQM[x] / (1000 ^ 2))),
                                   big.mark = ","
                                   ), 
                                 "km^2"
                                 ),
                               paste(
                                 "Hunting region: ",
                                 sp_wmus$HUNTING_REGION_ID[x],
                                 " (",
                                 sp_wmus$HUNTING_REGION_NAME[x],
                                 ")",
                                 sep = ''
                                 ),
                               # https://stackoverflow.com/a/29465942
                               paste(
                                 "Game management zone: ",
                                 sp_wmus$GAME_MANAGEMENT_ZONE_ID[x],
                                 " (",
                                 sp_wmus$GAME_MANAGEMENT_ZONE_NAME[x],
                                 ")",
                                 sep = ''
                               ),
                               sep = "<br/>"
                             )
                           })

# CREATE MAP
leaflet(sp_wmus) %>% 
  addTiles(tilesURL) %>%  
  PlotPolygons(., sp_wmus, "HUNTING_REGION_ID", wmu_hover_labels) %>% 
  AddStaticLabels(., wmu_static_labels, "WMU")

```

```{r}
# SHOW HUNTING REGIONS ON MAP

# PREPARE LABELS
# Labels to show on each hunting region all the time
# This also finds the centroid of each hunting region, to know where to place the labels.
hunting_region_static_labels <- gCentroid(sp_huntingRegions, byid = TRUE) %>% 
  as_tibble() %>% 
  cbind(seq(length(sp_huntingRegions)))
colnames(hunting_region_static_labels) <- c("long", "lat", "hunting_region")
# Labels to show on each hunting region when hovering
hunting_region_hover_labels <- lapply(seq(nrow(sp_huntingRegions)),
                           function(x) {
                             paste(
                               paste(
                                 "<b>Hunting region ", 
                                 sp_huntingRegions$HUNTING_REGION_ID[x], 
                                 " (",
                                 sp_huntingRegions$HUNTING_REGION_NAME[x],
                                 ")",
                                 "</b>", 
                                 sep = ''
                                 ),
                               paste(
                                 format(
                                   round(sp_huntingRegions$HUNTING_REGION_AREA[x]),
                                   big.mark = ",",
                                   ), 
                                 "km^2"
                                 ),
                               sep = "<br/>"
                             )
                           })

# Show on map
leaflet(sp_huntingRegions) %>% 
  addTiles(tilesURL) %>% 
  PlotPolygons(., sp_huntingRegions, "HUNTING_REGION_ID", hunting_region_hover_labels) %>% 
  AddStaticLabels(., hunting_region_static_labels, "hunting_region")
```

