---
title: "BC big game harvesting"
output: html_notebook
# output: 
#   flexdashboard::flex_dashboard
#   #flexdashboard::flex_dashboard:
#     #orientation: columns
---

<!-- 
Execute a chunk: click the *Run* button within the chunk or place your cursor inside it and press *Ctrl+Shift+Enter*.

Add a new chunk:  click the *Insert Chunk* button on the toolbar or press *Ctrl+Alt+I*.

This notebook analyzes data on big game harvesting in British Columbia, Canada.  It uses the 1976--2018 dataset from https://catalogue.data.gov.bc.ca/dataset/big-game-harvest-statistics-1976-2018  I used the .xlsx version of the dataset rather than the .csv version, because the csv had half the header text cut off and does not have descriptions of field codes.

Author:  Heather More
Started:  2020-11-24 

TODO:
http://www.env.gov.bc.ca/fw/wildlife/management-issues/docs/kootenay_mule_deer_faq.pdf
-->

# ```{r setup, include=FALSE}
```{r}
# Load packages
library(readxl) # read Excel spreadsheets
library(tidyverse) # data manipulation
library(janitor) # data cleaning
#library(leaflet)  # mapping
library(flexdashboard) # build output using knit button at top of RStudio

# Custom functions
source("GetColour.R")

# Define input parameters
# Location of data
dataPath  <- "D:/Documents/Code/BC big game harvesting/data"
fileName  <- "big-game-harvest-statistics-1976-2018.xlsx"
dataSheetName <- "BGHS"  # data
dataColumnTypes <- c("numeric", # hunt year
                     "text", # species
                     "text", # CI
                     "numeric", # prov flag
                     "text", # WMU
                     "text", # region
                     rep("numeric", 14)) # remaining columns
wmuName <- ""

# Which species to plot
speciesToShow <- c("ELK", "DEMU", "DEWT", "MOOS", "BEAB", "WOLF", "CARI", "GOAT", "SHEE", "BEAG", "COUG")

# Manually define species based on info in xlsx:  https://stackoverflow.com/a/10679467
# Get a list of all species codes by:  names(species_defs)
# Access a species name by:  species_defs[["CODE"]]
species_defs <- list(
  "ELK" = "Elk",
  "DEMU" = "Mule deer",
  "DEWT" = "White-tailed deer",
  "MOOS" = "Moose",
  "BEAB" = "Black bear",
  "WOLF" = "Wolf",
  "CARI" = "Caribou",
  "GOAT" = "Goat",
  "SHEE" = "Sheep",
  "BEAG" = "Grizzly bear",
  "COUG" = "Cougar"
)
# Function to look up full name of species based on code
GetSpeciesName <- function(speciesCode) {
  species_name <- species_defs[[speciesCode]]
  return(species_name)
}

# Colours for plotting
species_colours <- list(
  "MOOS" = "pink", # "Moose",
  "ELK" = "rust", # "Elk",
  "CARI" = "gold2", # "Caribou",
  "DEMU" = "yellow2", # "Mule deer",
  "DEWT" = "yellow", # "White-tailed deer",
  
  "GOAT" = "light green", # "Goat",
  "SHEE" = "dark green", # "Sheep",
  
  "BEAB" = "black", # "Black bear",
  "BEAG" = "grey", # "Grizzly bear",
  
  "WOLF" = "light blue", # "Wolf",
  
  "COUG" = "dark blue" # Cougar"
)
# Function to look up full name of species based on code
GetSpeciesColour <- function(speciesCode) {
  species_colour <- GetColour(species_colours[[speciesCode]])
  return(species_colour)
}

```

# ```{r, include=FALSE}
```{r}
# Read in data
df <- read_xlsx(file.path(dataPath, fileName, 
                           fsep = .Platform$file.sep), 
                sheet = dataSheetName,
                col_types = dataColumnTypes # specify type of data in each column, or it will use logical for some
                ) %>% 
  clean_names()  # make column names nice

# Investigate data
print(paste("Hunt years:  ", paste(min(df$hunt_year)), "--", paste(max(df$hunt_year)), collapse = ''))
print(paste('Species:  ', paste(lapply(unique(df$species), GetSpeciesName), collapse = ', ')))

# Find number of rows that report data for entire province
#unique(df$prov_flag)  # options are 0 or 9
df_province <- df %>% filter(prov_flag==9)
#unique(df_province$wmu)  # all wmu values are 999
```

```{r}
# Get data for plots
to_plot <- df_province %>% 
  filter(species %in% speciesToShow) %>% 
  rowwise() %>% # sum over each row:  https://stackoverflow.com/a/33806717
  mutate(
    total_kills = sum(resident_kills, non_resident_kills, na.rm=TRUE),
    total_hunters = sum(resident_hunters, non_resident_hunters, na.rm=TRUE),
    total_days = sum(resident_days, non_resident_days, na.rm=TRUE),
    kills_per_hunter = total_kills/total_hunters,
    days_per_hunter = total_days/total_hunters,
    days_per_kill = total_days/total_kills)
```

<!-- Column {data-width=333} -->
<!-- ------------------------------------- -->
```{r}
# Total kills
plot_prov_species <- ggplot(data = to_plot) +
  aes(x = hunt_year, y = total_kills, color = species) +
  geom_point(size=2, alpha=0.7) +
  geom_line() +
  theme_light() +
  scale_colour_manual(
    name = "Species",
    values = paste(sapply(names(species_colours), GetSpeciesColour)),
    breaks = names(species_colours),
    labels = paste(lapply(names(species_colours[]), GetSpeciesName))
  ) +
  labs(x = paste("Year"),
       y = paste("Total # kills"),
       title = paste("Total kills from", paste(min(to_plot$hunt_year)), "to", paste(max(to_plot$hunt_year))))
print(plot_prov_species)
```

<!-- Column {data-width=333} -->
<!-- ------------------------------------- -->
```{r}
# Total hunters
plot_prov_hunters <- ggplot(data = to_plot) +
  aes(x = hunt_year, y = total_hunters, color = species) +
  geom_point(size=2, alpha=0.7) +
  geom_line() +
  theme_light() +
  scale_colour_manual(
    name = "Species",
    values = paste(sapply(names(species_colours), GetSpeciesColour)),
    breaks = names(species_colours),
    labels = paste(lapply(names(species_colours[]), GetSpeciesName))
  ) +
  labs(x = paste("Year"),
       y = paste("Total # hunters"),
       title = paste("Total hunters from", paste(min(to_plot$hunt_year)), "to", paste(max(to_plot$hunt_year))))
print(plot_prov_hunters)
```

<!-- Column {data-width=333} -->
<!-- ------------------------------------- -->
```{r}
# Avg kills per hunter
plot_prov_killsperhunter <- ggplot(data = to_plot) +
  aes(x = hunt_year, y = kills_per_hunter, color = species) +
  geom_point(size=2, alpha=0.7) +
  geom_line() +
  theme_light() +
  scale_colour_manual(
    name = "Species",
    values = paste(sapply(names(species_colours), GetSpeciesColour)),
    breaks = names(species_colours),
    labels = paste(lapply(names(species_colours[]), GetSpeciesName))
  ) +
  labs(x = paste("Year"),
       y = paste("Avg kills per hunter"),
       title = paste("Average kills per hunter from", paste(min(to_plot$hunt_year)), "to", paste(max(to_plot$hunt_year))))
print(plot_prov_killsperhunter)
```
```{r}
# Total hunting days
plot_prov_totaldays <- ggplot(data = to_plot) +
  aes(x = hunt_year, y = total_days, color = species) +
  geom_point(size=2, alpha=0.7) +
  geom_line() +
  theme_light() +
  scale_colour_manual(
    name = "Species",
    values = paste(sapply(names(species_colours), GetSpeciesColour)),
    breaks = names(species_colours),
    labels = paste(lapply(names(species_colours[]), GetSpeciesName))
  ) +
  labs(x = paste("Year"),
       y = paste("Total # hunting days"),
       title = paste("Total # hunting days", paste(min(to_plot$hunt_year)), "to", paste(max(to_plot$hunt_year))))
print(plot_prov_totaldays)
```

```{r}
# Avg hunting days per hunter
plot_prov_daysperhunter <- ggplot(data = to_plot) +
  aes(x = hunt_year, y = days_per_hunter, color = species) +
  geom_point(size=2, alpha=0.7) +
  geom_line() +
  theme_light() +
  scale_colour_manual(
    name = "Species",
    values = paste(sapply(names(species_colours), GetSpeciesColour)),
    breaks = names(species_colours),
    labels = paste(lapply(names(species_colours[]), GetSpeciesName))
  ) +
  labs(x = paste("Year"),
       y = paste("Avg hunt days per hunter"),
       title = paste("Average # hunting days per hunter from", paste(min(to_plot$hunt_year)), "to", paste(max(to_plot$hunt_year))))
print(plot_prov_daysperhunter)
```

```{r}
# Avg hunting days per kill
plot_prov_daysperhunter <- ggplot(data = to_plot) +
  aes(x = hunt_year, y = days_per_kill, color = species) +
  geom_point(size=2, alpha=0.7) +
  geom_line() +
  theme_light() +
  scale_colour_manual(
    name = "Species",
    values = paste(sapply(names(species_colours), GetSpeciesColour)),
    breaks = names(species_colours),
    labels = paste(lapply(names(species_colours[]), GetSpeciesName))
  ) +
  labs(x = paste("Year"),
       y = paste("Avg hunt days per kill"),
       title = paste("Average # hunting days per kill from", paste(min(to_plot$hunt_year)), "to", paste(max(to_plot$hunt_year))))
print(plot_prov_daysperhunter)
```

```{r}
# Resident and non-resident hunters
# Note that each individual hunter is probably counted multiple times, because a lot of them probably have more than one tag
hunters_residence <- df_province %>% 
  group_by(hunt_year) %>% 
  summarise(
    across(resident_hunters, sum, na.rm=TRUE),
    across(non_resident_hunters, sum, na.rm=TRUE),
  ) %>% 
  pivot_longer(cols = !hunt_year, names_to = "residency", values_to = "n_hunters")

plot_prov_resnonres <- ggplot(data = hunters_residence) +
  aes(x = hunt_year, y = n_hunters, color = residency) +
  geom_point(size=2, alpha=0.7) +
  geom_line() +
  theme_light() +
  labs(x = paste("Year"),
       y = paste("# hunters"),
       title = paste("Resident and non-resident hunters", paste(min(to_plot$hunt_year)), "to", paste(max(to_plot$hunt_year))))
print(plot_prov_resnonres)
```

