---
title: "BC big game harvesting"
output: html_notebook
# output: 
#   flexdashboard::flex_dashboard
#   #flexdashboard::flex_dashboard:
#     #orientation: columns
---

<!-- 
Execute a chunk: click the *Run* button within the chunk or place your cursor inside it and press *Ctrl+Shift+Enter*.

Add a new chunk:  click the *Insert Chunk* button on the toolbar or press *Ctrl+Alt+I*.

This notebook analyzes data on big game harvesting in British Columbia, Canada.  It uses the 1976--2018 dataset from https://catalogue.data.gov.bc.ca/dataset/big-game-harvest-statistics-1976-2018  I used the .xlsx version of the dataset rather than the .csv version, because the csv had half the header text cut off and does not have descriptions of field codes.

Annual population of BC from:  https://www2.gov.bc.ca/gov/content/data/statistics/people-population-community/population/population-estimates

Hunting licence sales from:
https://catalogue.data.gov.bc.ca/dataset/hunting-sales-statistics-2005-2019

BC Wildlife management units from:
https://catalogue.data.gov.bc.ca/dataset/wildlife-management-units

Author:  Heather More
Started:  2020-11-24 

TODO:
http://www.env.gov.bc.ca/fw/wildlife/management-issues/docs/kootenay_mule_deer_faq.pdf
try getting population data from here:  http://a100.gov.bc.ca/pub/siwe/search_reset.do
BC hunter expenditure survey:  http://www.env.gov.bc.ca/fw/wildlife/bc_hunt_value.html
BC harvest allocation policies:  http://www.env.gov.bc.ca/fw/wildlife/harvest_alloc/
region and wmu info:  https://www2.gov.bc.ca/gov/content/sports-culture/recreation/fishing-hunting/hunting/regulations-synopsis
BC regions map:  https://governmentofbc.maps.arcgis.com/home/webmap/viewer.html
BC data sets:  catalogue.data.gov.bc.ca/dataset
BC game management zones:  https://catalogue.data.gov.bc.ca/dataset/game-management-zones
-->

# ```{r setup, include=FALSE}
```{r}
# Load packages
library(readxl) # read Excel spreadsheets
library(tidyverse) # data manipulation
library(janitor) # data cleaning
library(leaflet)  # mapping
library(geojsonio)  # read geojson files
library(rmapshaper)  # simplify areas on maps
library(rgeos)  # functions for handling geographic regions
library(RColorBrewer)  # colour schemes
#library(pgirmess)  # convert geojson to gpx
#library(sp)  # spatial polygon analysis
library(flexdashboard) # build output using knit button at top of RStudio
library(htmltools)  # HTML formatting for map labels

# Custom functions
source("GetColour.R")

# Define input parameters
# Location of data
dataPath  <- "D:/Documents/Code/BC big game harvesting/data"
harvestName  <- "big-game-harvest-statistics-1976-2018.xlsx"
dataSheetName <- "BGHS"  # data
dataColumnTypes <- c("numeric", # hunt year
                     "text", # species
                     "text", # CI
                     "numeric", # prov flag
                     "text", # WMU
                     "text", # region
                     rep("numeric", 14)) # remaining columns
bcPopName <- "pop_bc_annual_estimates_clean.xlsx"
popDataSheetName <- "data"
wmuName <- "wmu-map/WAA_WILDLIFE_MGMT_UNITS_SVW.geojson"
tilesURL <- "http://server.arcgisonline.com/ArcGIS/rest/services/Canvas/World_Light_Gray_Base/MapServer/tile/{z}/{y}/{x}"

# Which species to plot
speciesToShow <- c("ELK", "DEMU", "DEWT", "MOOS", "BEAB", "WOLF", "CARI", "GOAT", "SHEE", "BEAG", "COUG")

# Manually define species based on info in xlsx:  https://stackoverflow.com/a/10679467
# Get a list of all species codes by:  names(species_defs)
# Access a species name by:  species_defs[["CODE"]]
species_defs <- list(
  "ELK" = "Elk",
  "DEMU" = "Mule deer",
  "DEWT" = "White-tailed deer",
  "MOOS" = "Moose",
  "BEAB" = "Black bear",
  "WOLF" = "Wolf",
  "CARI" = "Caribou",
  "GOAT" = "Goat",
  "SHEE" = "Sheep",
  "BEAG" = "Grizzly bear",
  "COUG" = "Cougar"
)
# Function to look up full name of species based on code
GetSpeciesName <- function(speciesCode) {
  species_name <- species_defs[[speciesCode]]
  return(species_name)
}

# Colours for plotting
species_colours <- list(
  "MOOS" = "pink", # "Moose",
  "ELK" = "rust", # "Elk",
  "CARI" = "gold2", # "Caribou",
  "DEMU" = "yellow2", # "Mule deer",
  "DEWT" = "yellow", # "White-tailed deer",
  
  "GOAT" = "light green", # "Goat",
  "SHEE" = "dark green", # "Sheep",
  
  "BEAB" = "black", # "Black bear",
  "BEAG" = "grey", # "Grizzly bear",
  
  "WOLF" = "light blue", # "Wolf",
  
  "COUG" = "dark blue" # Cougar"
)
# Function to look up full name of species based on code
GetSpeciesColour <- function(speciesCode) {
  species_colour <- GetColour(species_colours[[speciesCode]])
  return(species_colour)
}

```

# ```{r, include=FALSE}
```{r}
# Read in data
df <- read_xlsx(file.path(dataPath, harvestName, 
                           fsep = .Platform$file.sep), 
                sheet = dataSheetName,
                col_types = dataColumnTypes # specify type of data in each column, or it will use logical for some
                ) %>% 
  clean_names()  # make column names nice

df_pop <- read_xlsx(file.path(dataPath, bcPopName, 
                           fsep = .Platform$file.sep),
                    sheet = popDataSheetName) %>% 
  clean_names()  # make column names nice

# Investigate data
print(paste("Hunt years:  ", paste(min(df$hunt_year)), "--", paste(max(df$hunt_year)), collapse = ''))
print(paste('Species:  ', paste(lapply(unique(df$species), GetSpeciesName), collapse = ', ')))
print(paste('WMUs:  ', paste(length(unique(df$wmu)))))
print(paste('Regions:  ', paste(length(unique(df$region)))))

# Find number of rows that report data for entire province
#unique(df$prov_flag)  # options are 0 or 9
df_province <- df %>% filter(prov_flag==9)
#unique(df_province$wmu)  # all wmu values are 999
```

```{r}
# Get data for plots
to_plot <- df_province %>% 
  filter(species %in% speciesToShow) %>% 
  rowwise() %>% # sum over each row:  https://stackoverflow.com/a/33806717
  mutate(
    total_kills = sum(resident_kills, non_resident_kills, na.rm=TRUE),
    total_hunters = sum(resident_hunters, non_resident_hunters, na.rm=TRUE),
    total_days = sum(resident_days, non_resident_days, na.rm=TRUE),
    kills_per_hunter = total_kills/total_hunters,
    days_per_hunter = total_days/total_hunters,
    days_per_kill = total_days/total_kills)

df_pop_years <- df_pop %>% 
  filter(year >= min(to_plot$hunt_year) 
         & year <= max(to_plot$hunt_year)) %>% 
  rename(population_human = population)
to_plot <- to_plot %>% 
  left_join(df_pop_years, by = c("hunt_year" = "year"))  # https://dplyr.tidyverse.org/articles/two-table.html


```

<!-- Column {data-width=333} -->
<!-- ------------------------------------- -->
```{r}
# Total kills
plot_prov_species <- ggplot(data = to_plot) +
  aes(x = hunt_year, y = total_kills, color = species) +
  geom_point(size=2, alpha=0.7) +
  geom_line() +
  theme_light() +
  scale_colour_manual(
    name = "Species",
    values = paste(sapply(names(species_colours), GetSpeciesColour)),
    breaks = names(species_colours),
    labels = paste(lapply(names(species_colours[]), GetSpeciesName))
  ) +
  labs(x = paste("Year"),
       y = paste("Total # kills"),
       title = paste("Total kills from", paste(min(to_plot$hunt_year)), "to", paste(max(to_plot$hunt_year))))
print(plot_prov_species)
```

<!-- Column {data-width=333} -->
<!-- ------------------------------------- -->
```{r}
# Total hunters
plot_prov_hunters <- ggplot(data = to_plot) +
  aes(x = hunt_year, y = total_hunters, color = species) +
  geom_point(size=2, alpha=0.7) +
  geom_line() +
  theme_light() +
  scale_colour_manual(
    name = "Species",
    values = paste(sapply(names(species_colours), GetSpeciesColour)),
    breaks = names(species_colours),
    labels = paste(lapply(names(species_colours[]), GetSpeciesName))
  ) +
  labs(x = paste("Year"),
       y = paste("Total # hunters"),
       title = paste("Total hunters from", paste(min(to_plot$hunt_year)), "to", paste(max(to_plot$hunt_year))))
print(plot_prov_hunters)
```

<!-- Column {data-width=333} -->
<!-- ------------------------------------- -->
```{r}
# Avg kills per hunter
plot_prov_killsperhunter <- ggplot(data = to_plot) +
  aes(x = hunt_year, y = kills_per_hunter, color = species) +
  geom_point(size=2, alpha=0.7) +
  geom_line() +
  theme_light() +
  scale_colour_manual(
    name = "Species",
    values = paste(sapply(names(species_colours), GetSpeciesColour)),
    breaks = names(species_colours),
    labels = paste(lapply(names(species_colours[]), GetSpeciesName))
  ) +
  labs(x = paste("Year"),
       y = paste("Avg kills per hunter"),
       title = paste("Average kills per hunter from", paste(min(to_plot$hunt_year)), "to", paste(max(to_plot$hunt_year))))
print(plot_prov_killsperhunter)
```
```{r}
# Total hunting days
plot_prov_totaldays <- ggplot(data = to_plot) +
  aes(x = hunt_year, y = total_days, color = species) +
  geom_point(size=2, alpha=0.7) +
  geom_line() +
  theme_light() +
  scale_colour_manual(
    name = "Species",
    values = paste(sapply(names(species_colours), GetSpeciesColour)),
    breaks = names(species_colours),
    labels = paste(lapply(names(species_colours[]), GetSpeciesName))
  ) +
  labs(x = paste("Year"),
       y = paste("Total # hunting days"),
       title = paste("Total # hunting days", paste(min(to_plot$hunt_year)), "to", paste(max(to_plot$hunt_year))))
print(plot_prov_totaldays)
```

```{r}
# Avg hunting days per hunter
plot_prov_daysperhunter <- ggplot(data = to_plot) +
  aes(x = hunt_year, y = days_per_hunter, color = species) +
  geom_point(size=2, alpha=0.7) +
  geom_line() +
  theme_light() +
  scale_colour_manual(
    name = "Species",
    values = paste(sapply(names(species_colours), GetSpeciesColour)),
    breaks = names(species_colours),
    labels = paste(lapply(names(species_colours[]), GetSpeciesName))
  ) +
  labs(x = paste("Year"),
       y = paste("Avg hunt days per hunter"),
       title = paste("Average # hunting days per hunter from", paste(min(to_plot$hunt_year)), "to", paste(max(to_plot$hunt_year))))
print(plot_prov_daysperhunter)
```

```{r}
# Avg hunting days per kill
plot_prov_daysperhunter <- ggplot(data = to_plot) +
  aes(x = hunt_year, y = days_per_kill, color = species) +
  geom_point(size=2, alpha=0.7) +
  geom_line() +
  theme_light() +
  scale_colour_manual(
    name = "Species",
    values = paste(sapply(names(species_colours), GetSpeciesColour)),
    breaks = names(species_colours),
    labels = paste(lapply(names(species_colours[]), GetSpeciesName))
  ) +
  labs(x = paste("Year"),
       y = paste("Avg hunt days per kill"),
       title = paste("Average # hunting days per kill from", paste(min(to_plot$hunt_year)), "to", paste(max(to_plot$hunt_year))))
print(plot_prov_daysperhunter)
```



```{r}
# Resident hunters of each species
plot_prov_speciesres <- ggplot(data = to_plot) +
  aes(x = hunt_year, y = resident_hunters, color = species) +
  geom_point(size=2, alpha=0.7) +
  geom_line() +
  theme_light() +
  scale_colour_manual(
    name = "Species",
    values = paste(sapply(names(species_colours), GetSpeciesColour)),
    breaks = names(species_colours),
    labels = paste(lapply(names(species_colours[]), GetSpeciesName))
  ) +
  labs(x = paste("Year"),
       y = paste("# resident hunters"),
       title = paste("Resident hunters from", paste(min(to_plot$hunt_year)), "to", paste(max(to_plot$hunt_year))))
print(plot_prov_speciesres)
```

```{r}
# Non-resident hunters of each species
plot_prov_speciesnonres <- ggplot(data = to_plot) +
  aes(x = hunt_year, y = non_resident_hunters, color = species) +
  geom_point(size=2, alpha=0.7) +
  geom_line() +
  theme_light() +
  scale_colour_manual(
    name = "Species",
    values = paste(sapply(names(species_colours), GetSpeciesColour)),
    breaks = names(species_colours),
    labels = paste(lapply(names(species_colours[]), GetSpeciesName))
  ) +
  labs(x = paste("Year"),
       y = paste("# non-resident hunters"),
       title = paste("Non-resident hunters from", paste(min(to_plot$hunt_year)), "to", paste(max(to_plot$hunt_year))))
print(plot_prov_speciesnonres)
```

```{r}
# Resident hunters of each species, normalized to BC population
plot_prov_speciesresfrac <- ggplot(data = to_plot) +
  aes(x = hunt_year, y = resident_hunters/population_human, color = species) +
  geom_point(size=2, alpha=0.7) +
  geom_line() +
  theme_light() +
  scale_colour_manual(
    name = "Species",
    values = paste(sapply(names(species_colours), GetSpeciesColour)),
    breaks = names(species_colours),
    labels = paste(lapply(names(species_colours[]), GetSpeciesName))
  ) +
  labs(x = paste("Year"),
       y = paste("# resident hunters/total BC population"),
       title = paste("Resident hunters as fraction of BC population from", paste(min(to_plot$hunt_year)), "to", paste(max(to_plot$hunt_year))))
print(plot_prov_speciesresfrac)
```

```{r}
# Total resident and non-resident hunters
# Note that each individual hunter is probably counted multiple times, because a lot of them probably have more than one tag
hunters_residence <- df_province %>% 
  group_by(hunt_year) %>% 
  summarise(
    across(resident_hunters, sum, na.rm=TRUE),
    across(non_resident_hunters, sum, na.rm=TRUE),
  ) %>% 
  pivot_longer(cols = !hunt_year, names_to = "residency", values_to = "n_hunters") %>% 
  left_join(df_pop_years, by = c("hunt_year" = "year"))

plot_prov_totalresnonres <- ggplot(data = hunters_residence) +
  aes(x = hunt_year, y = n_hunters, color = residency) +
  geom_point(size=2, alpha=0.7) +
  geom_line() +
  theme_light() +
  labs(x = paste("Year"),
       y = paste("# hunters"),
       title = paste("Total resident and non-resident hunters", paste(min(to_plot$hunt_year)), "to", paste(max(to_plot$hunt_year))))
print(plot_prov_totalresnonres)
```



```{r}
plot_prov_totalresfrac <- ggplot(data = filter(hunters_residence, residency == "resident_hunters")) +
  aes(x = hunt_year, y = n_hunters/population_human) +
  geom_point(size=2, alpha=0.7) +
  geom_line() +
  theme_light() +
  labs(x = paste("Year"),
       y = paste("# hunters/BC population"),
       title = paste("Total resident hunters as a fraction of BC population", paste(min(to_plot$hunt_year)), "to", paste(max(to_plot$hunt_year))))
print(plot_prov_totalresfrac)
```

```{r}
plot_bcpop <- ggplot(data = df_pop_years) +
  aes(x = year, y = population_human) +
  geom_point(size=2, alpha=0.7) +
  geom_line() +
  theme_light() +
  labs(x = paste("Year"),
       y = paste("BC population"),
       title = paste("BC population", paste(min(to_plot$hunt_year)), "to", paste(max(to_plot$hunt_year))))
print(plot_bcpop)
```


```{r}
# Avg kills per hunter, separated by hunter residency
plot_prov_killsperhunter_resident <- ggplot(data = to_plot) +
  aes(x = hunt_year, y = resident_kills/resident_hunters, color = species) +
  geom_point(size=2, alpha=0.7) +
  geom_line() +
  theme_light() +
  scale_colour_manual(
    name = "Species",
    values = paste(sapply(names(species_colours), GetSpeciesColour)),
    breaks = names(species_colours),
    labels = paste(lapply(names(species_colours[]), GetSpeciesName))
  ) +
  labs(x = paste("Year"),
       y = paste("Avg kills per resident hunter"),
       title = paste("Average kills per resident hunter from", paste(min(to_plot$hunt_year)), "to", paste(max(to_plot$hunt_year))))
print(plot_prov_killsperhunter_resident)
```

```{r}
plot_prov_killsperhunter_nonresident <- ggplot(data = to_plot) +
  aes(x = hunt_year, y = non_resident_kills/non_resident_hunters, color = species) +
  geom_point(size=2, alpha=0.7) +
  geom_line() +
  theme_light() +
  scale_colour_manual(
    name = "Species",
    values = paste(sapply(names(species_colours), GetSpeciesColour)),
    breaks = names(species_colours),
    labels = paste(lapply(names(species_colours[]), GetSpeciesName))
  ) +
  labs(x = paste("Year"),
       y = paste("Avg kills per non-resident hunter"),
       title = paste("Average kills per non-resident hunter from", paste(min(to_plot$hunt_year)), "to", paste(max(to_plot$hunt_year))))
print(plot_prov_killsperhunter_nonresident)
```

```{r}
# Avg hunting days per kill, separated by hunter residency
plot_prov_daysperhunter_resident <- ggplot(data = to_plot) +
  aes(x = hunt_year, y = resident_days/resident_kills, color = species) +
  geom_point(size=2, alpha=0.7) +
  geom_line() +
  theme_light() +
  scale_colour_manual(
    name = "Species",
    values = paste(sapply(names(species_colours), GetSpeciesColour)),
    breaks = names(species_colours),
    labels = paste(lapply(names(species_colours[]), GetSpeciesName))
  ) +
  labs(x = paste("Year"),
       y = paste("Avg hunt days per kill"),
       title = paste("Average # hunting days per kill for resident hunters from", paste(min(to_plot$hunt_year)), "to", paste(max(to_plot$hunt_year))))
print(plot_prov_daysperhunter_resident)
```

```{r}
plot_prov_daysperhunter_nonresident <- ggplot(data = to_plot) +
  aes(x = hunt_year, y = non_resident_days/non_resident_kills, color = species) +
  geom_point(size=2, alpha=0.7) +
  geom_line() +
  theme_light() +
  scale_colour_manual(
    name = "Species",
    values = paste(sapply(names(species_colours), GetSpeciesColour)),
    breaks = names(species_colours),
    labels = paste(lapply(names(species_colours[]), GetSpeciesName))
  ) +
  labs(x = paste("Year"),
       y = paste("Avg hunt days per kill"),
       title = paste("Average # hunting days per kill for non-resident hunters from", paste(min(to_plot$hunt_year)), "to", paste(max(to_plot$hunt_year))))
print(plot_prov_daysperhunter_nonresident)
```

```{r}
total_allspecies <- to_plot %>% 
  group_by(hunt_year) %>% 
  summarise(
    across(resident_kills, sum, na.rm=TRUE),
    across(non_resident_kills, sum, na.rm=TRUE),
  ) %>% 
  pivot_longer(cols = !hunt_year, names_to = "residency", values_to = "n_kills")

plot_prov_totalkills <- ggplot(data = total_allspecies) +
  aes(x = hunt_year, y = n_kills, color = residency) +
  geom_point(size=2, alpha=0.7) +
  geom_line() +
  theme_light() +
  labs(x = paste("Year"),
       y = paste("Total kills"),
       title = paste("Total kills over all species from", paste(min(to_plot$hunt_year)), "to", paste(max(to_plot$hunt_year))))
print(plot_prov_totalkills)
```

```{r, eval = FALSE}
# Convert Jim's trapline data
jim_trapline <- geojsonio::geojson_read(file.path(dataPath, "jim-trapline-map/WAA_TRAPLINE_AREAS_SP.geojson", 
                           fsep = .Platform$file.sep), what = "sp")
trapline_id = "TR0332T008"
trapline_ind <- which(sapply(jim_trapline$TRAPLINE_AREA_IDENTIFIER, FUN = function(x) trapline_id %in% x))
trapline_ind <- trapline_ind[[1]]
trapline_df <- fortify(jim_trapline) %>% 
  filter(id==trapline_ind) %>% 
  select(id, long, lat)
writeGPX(trapline_df, paste(trapline_id, "_boundary"), type = "t")  # https://rdrr.io/cran/pgirmess/man/writeGPX.html
```

```{r}
# Wildlife management units map
# https://rstudio.github.io/leaflet/json.html
# 4,535,784 bytes
wmus <- geojsonio::geojson_read(file.path(dataPath, wmuName, 
                           fsep = .Platform$file.sep), what = "sp")
# Simplify boundaries so they render faster
# 963,312 bytes
wmus <- rmapshaper::ms_simplify(wmus)
# Check total areas add to area of BC (944,735 km^2)
# Note that the geojson areas include a bunch of water so we expect a higher estimate
#testarea <- sum(wmus$FEATURE_AREA_SQM/(1000^2))
# Show possible colour schemes
#display.brewer.all()
# Generate palette
mapColours <- colorFactor(
  palette = brewer.pal(length(unique(wmus$REGION_RESPONSIBLE)),
                         "Set2"),
  levels = unique(wmus$REGION_RESPONSIBLE)
  )
# Prepare labels
#https://stackoverflow.com/a/43155126
wmu_hover_labels <- lapply(seq(nrow(wmus)), function(x) {
  paste(
           paste("<b>WMU # ", wmus$WILDLIFE_MGMT_UNIT_ID[x], "</b>", sep = ''),
           paste(format(
             round(as.numeric(wmus$FEATURE_AREA_SQM[x]/(1000^2)), 1),
             nsmall = 1,
             big.mark = ","
             ), "km^2"),  # https://stackoverflow.com/a/29465942
           paste("Game management zone: ", wmus$GAME_MANAGEMENT_ZONE_ID[x],
                 " (", wmus$GAME_MANAGEMENT_ZONE_NAME[x], ")",
                 sep = ''),
           paste("Hunting region: ", wmus$REGION_RESPONSIBLE[x],
                 " (", wmus$REGION_RESPONSIBLE_NAME[x], ")",
                 sep = ''),
         sep = "<br/>"
  )
})
wmu_static_labels <- lapply(wmus@polygons, function(x) {x@labpt}) %>% 
  unlist() %>% 
  matrix(., nrow = length(.), ncol = 2, byrow = TRUE) %>% 
  as_tibble() %>% 
  cbind(wmus$WILDLIFE_MGMT_UNIT_ID)
colnames(wmu_static_labels) <- c("long", "lat", "WMU")
  
# Show on map
leaflet(wmus) %>% 
  addTiles(tilesURL) %>% 
  addPolygons(stroke = TRUE,
              weight = 1,
              opacity = 1,
              fillOpacity = 0.4,
              color = ~mapColours(REGION_RESPONSIBLE),
              label = lapply(wmu_hover_labels, htmltools::HTML),  # labels on hover:  https://stackoverflow.com/a/43155126
  highlightOptions = highlightOptions(color = "red",
                                      fillColor = "red",
                                      weight = 2,
                                      bringToFront = TRUE)  # bring hovered-over polygon to front
  ) %>% 
  addLabelOnlyMarkers(data = wmu_static_labels,  # labels always on polygons
                      lng = ~long, 
                      lat = ~lat, 
                      label = ~WMU,
                      labelOptions = labelOptions(
                        noHide = TRUE, 
                        direction = 'auto', 
                        textOnly = TRUE,
                        style = list(
                          "color" = "grey"
                          )
                        )
                      )
```

```{r}
# Show hunting regions on map
# Get list of WMUs mapped to hunting regions
# WMUs ending in 00 denote values from that hunting region but unknown WMU
# WMUs ending in 99 denote totals for that hunting region
# WMU 999 is province-wide total
# wmu_regions <- df %>% 
#   select(wmu, region) %>% 
#   distinct()
# WMU codes in geojson are e.g. 1-1, 1-15, but in hunting data they are e.g. 101, 115
# regions <- wmu_regions[wmus$WILDLIFE_MGMT_UNIT_ID]

# https://gis.stackexchange.com/a/225731
#region1 <- gUnaryUnion(wmus[wmus@data$REGION_RESPONSIBLE == 1, ])
hunting_regions <- gUnaryUnion(wmus, id = wmus@data$REGION_RESPONSIBLE)
#plot(hunting_regions)
# Access area of each region
# UPDATE:  note that "area" in a polygon class is NOT the projected area, but rather a planar area measurement used to ensure smaller polygons are plotted after larger ones:  https://www.rdocumentation.org/packages/sp/versions/1.4-4/topics/Polygons-class
# GetArea <- function(x) {
#   x@area
# }
# hunting_region_areas<-sapply(hunting_regions@polygons, GetArea)
# Show on map
leaflet(hunting_regions) %>% 
  addTiles(tilesURL) %>% 
  addPolygons(stroke = TRUE,
              weight = 1,
              opacity = 1,
              fillOpacity = 0.4,
              #color = ~mapColours(id),
              # popup = ~paste(sep = "<br/>",
              #                paste("<b>Hunting region: ", ID, 
              #                      " (", ID, ")", "</b>",
              #                      sep = '')
              # ),
              highlightOptions = highlightOptions(color = "red",
                                                  fillColor = "red",
                                                  weight = 2,
                                                  bringToFront = TRUE)  # bring hovered-over polygon to front
              )
```

